Functional oxide layers for electrical isolation and chemical passivation of steel substrates

### 2023-02-13 Thesis/chapter/033-linreg-krr-svm.tex
start kernel

%\textbf{Kernel Ridge Regression}
\textbf{\gls{krr}} combines ridge regression with the kernel method. 
A~kernel transforms data in such a way that a linear hyperplane (a line in one dimension, 
a plane in two dimensions) can fit data in regression problems or seperate the data in classification problems. 
A kernel is some kind of similarity measure 
which fulfills the requirements of nonnegativity, symmetry, linearity\cite{rupp2015machine}.
The following equations show definitions for 
linear (eq.~\ref{eq:lin-kernel}), 
polynomial (eq.~\ref{eq:pol-kernel}), 
sigmoidal (eq.~\ref{eq:sig-kernel}) and 
radial basis functionial kernel~(eq. \ref{eq:rbf-kernel}), with $\gamma$ as fixed hyperparameter and $c_0$ as parameter to optimize.
\begin{align}
    \label{eq:lin-kernel}
    k_{lin}(\mathbf{x},\mathbf{y}) &= \mathbf{x}^{\top} \mathbf{y} \\
    \label{eq:pol-kernel}
    k_{pol}(\mathbf{x}, \mathbf{y}) &= (\gamma \mathbf{x}^{\top} \mathbf{y} + c_0)^d \\
    \label{eq:sig-kernel}
    k_{sig}(\mathbf{x}, \mathbf{y}) &= \text{tanh}(\gamma \mathbf{x}^{\top} \mathbf{y} + c_0) \\
    \label{eq:rbf-kernel}
    k_{rbf}(\mathbf{x}, \mathbf{y}) &= \text{exp}(- \gamma \| \mathbf{x} -  \mathbf{y} \|^2) 
\end{align}

%\textbf{Support vector machines} 
\textbf{\Gls{svm}} are versatile machine learning algorithm first mentioned in 1992\cite{boser1992training}. 
%
The \gls{svm} was initially developed by V. Vapnik for the binary classification of seperable data, then improved to handle non-sperable data and eventually adapted to solve regression problems.
The concepts of \gls{svm} will be discussed in the same chronological order. 
%
Classifiction works by spanning a hyperplane between two linearly separable categories in a way such that the closest points from each category have the largest distance to the hyperplane. 
The distance from the points to the hyperplane is called margin $\tau$.
These points are called support vectors and are used to define the hyperplane.
A \gls{svm} avoids overfitting by only using a subset of the data - the support vectors - to fit the model. 
The goal is to find the decision boundary which correctly classifies all samples with the biggest margin. 
%
The decision boundary can be expressed as an hyperplane
\begin{equation}
	\hat{\mathbf{y}} = h(\mathbf{x}) = \mathbf{w} \cdot \mathbf{x} + b
\end{equation}
with $
\mathbf{y} \in \lbrace +1, -1 \rbrace^n ,\,
\hat{\mathbf{y}} \in \mathbb{R}^n ,\,
\mathbf{w} \in \mathbb{R}^d ,\,
\mathbf{x} \in \mathbb{R}^{n \times d} 
$.
The constraint of the positive and negative support vectors ($x^+$ and $x^-$, respectively) satisfying
\begin{align}
	\mathbf{w} \cdot x^+ + b &= 1 \\
	\mathbf{w} \cdot x^- + b &= -1 
\end{align}
can be generalized to 
\begin{equation}
	\label{eq:svm01}
	y_i(\mathbf{w} \cdot \mathbf{x}_i + b ) \geqslant 1 %\qquad i=1,\dots,n
\end{equation}
where $y_i$ being the labels of the training data. 
The width of the margin can be inferred by projecting the vector spanning between two support vectors 
on opposite sides of the decision boundary onto the unit vector perpendicular to the hyperplane. 
If we now take a vector from a positive support vector $x^+$ to a negative support vector $x^-$ and project it on to the unit vector of $w$ (which is perpendicular to the hyperplane), 
we get the width of the margin. 
\iffalse
\begin{equation}
	\label{eq:svm02}
	\begin{aligned} 
		2\tau &= (x^+ - x^-) \cdot \frac{\mathbf{w}}{\|\mathbf{w}\|} \\
			&= (x^+ \cdot \mathbf{w} - x^-\cdot \mathbf{w} ) \cdot \frac{1}{\|\mathbf{w}\|} \\
			&= ( (1-b) - (-1-b) ) \cdot \frac{1}{\|\mathbf{w}\|} \\
			&= \frac{2}{\|w\|}
	\end{aligned}
\end{equation}
\fi
\begin{equation}
	\label{eq:svm02}
	\begin{aligned} 
		2\tau &= (x^+ - x^-) \cdot \frac{\mathbf{w}}{\|\mathbf{w}\|} 
			= (x^+ \cdot \mathbf{w} - x^-\cdot \mathbf{w} ) \cdot \frac{1}{\|\mathbf{w}\|} \\
			&= ( (1-b) - (-1-b) ) \cdot \frac{1}{\|\mathbf{w}\|} 
			= \frac{2}{\|\mathbf{w}\|}
	\end{aligned}
\end{equation}
%\td{draw sketch and make \ref{eq:svm02} on two lines}
Thus maximizing the margin is equivalent with minimizing $\|\mathbf{w}\|$ and minimizing $\frac{1}{2}\|\mathbf{w}\|^2$.
By incorporation the contraint (eq. \ref{eq:svm01}) via the Lagrangian multiplier method we arrive at the loss function: 
\begin{equation}
	\label{eq:svm03}
	\mathcal{L} = \frac{1}{2} \|\mathbf{w}\| - \sum_i^n \alpha_i [ y_i ( w_i \cdot x_i + b) -1 ] 
\end{equation}
which can be rewritten by some mathematical wizardry (setting the partial derivatives of the Lagrangian function
$\frac{\partial \mathcal{L}}{\partial \mathbf{w}}$ and $\frac{\partial \mathcal{L}}{\partial b}$ 
to zero and inserting into eq. (\ref{eq:svm03}))\cite{winston1992artificial,cherkassky1998learning}: 
\begin{equation}
	\mathcal{L} = \sum_i^n \alpha_i - \frac{1}{2} \sum_i^n \sum_j^n \alpha_i \alpha_j y_i y_j \mathbf{x}_i \cdot \mathbf{x}_j 
\end{equation}
%
\td{write about kernel trick and non strick}

>>>>

%\textbf{Kernel Ridge Regression}
%%%%%%%%%%%%%%%%%%%%%%%%%%% { Kernel Ridge Regression } KRR %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{\gls{krr}} combines ridge regression with the kernel method. 
A~kernel transforms data in such a way that a linear hyperplane (a line in one dimension, 
a plane in two dimensions) can fit data in regression problems or seperate the data in classification problems. 
A kernel is some kind of similarity measure 
which fulfills the requirements of nonnegativity, symmetry, linearity\cite{rupp2015machine}.
The following equations show definitions for 
linear (eq.~\ref{eq:lin-kernel}), 
polynomial (eq.~\ref{eq:pol-kernel}), 
sigmoidal (eq.~\ref{eq:sig-kernel}) and 
radial basis functionial kernel~(eq. \ref{eq:rbf-kernel}), with $\gamma$ as fixed hyperparameter and $c_0$ as parameter to optimize.
\begin{align}
    \label{eq:lin-kernel}
    k_{lin}(\mathbf{x},\mathbf{y}) &= \mathbf{x}^{\top} \mathbf{y} \\
    \label{eq:pol-kernel}
    k_{pol}(\mathbf{x}, \mathbf{y}) &= (\gamma \mathbf{x}^{\top} \mathbf{y} + c_0)^d \\
    \label{eq:sig-kernel}
    k_{sig}(\mathbf{x}, \mathbf{y}) &= \text{tanh}(\gamma \mathbf{x}^{\top} \mathbf{y} + c_0) \\
    \label{eq:rbf-kernel}
    k_{rbf}(\mathbf{x}, \mathbf{y}) &= \text{exp}(- \gamma \| \mathbf{x} -  \mathbf{y} \|^2) 
\end{align}
A requirement for using kernels is having a dot product in the loss function. 
This can be accomplished by expressing $\mathbf{k}$ with $\mathbf{X}$: $\mathbf{k}=\mathbf{X}^\top \mathbf{r}$
\cite{rudin2020least}

%%%%%%%%%%%%%%%%%%%%%%%%%%% { Support vector machines } SVM %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textbf{\Gls{svm}} are versatile machine learning algorithm first mentioned in 1992\cite{boser1992training}. 
%
The \gls{svm} was initially developed by V. Vapnik for the binary classification of seperable data, then improved to handle non-sperable data and eventually adapted to solve regression problems.
The concepts of \gls{svm} will be discussed in the same chronological order. 
%
Classifiction works by spanning a hyperplane between two linearly separable categories in a way such that the closest points from each category have the largest distance to the hyperplane. 
The distance from the points to the hyperplane is called margin $\tau$.
These points are called support vectors and are used to define the hyperplane.
A \gls{svm} avoids overfitting by only using a subset of the data - the support vectors - to fit the model. 
The goal is to find the decision boundary which correctly classifies all samples with the biggest margin. 
%
The decision boundary can be expressed as an hyperplane
\begin{equation}
	\hat{\mathbf{y}} = h(\mathbf{x}) = \mathbf{w} \cdot \mathbf{x} + b
\end{equation}
with $
\mathbf{y} \in \lbrace +1, -1 \rbrace^n ,\,
\hat{\mathbf{y}} \in \mathbb{R}^n ,\,
\mathbf{w} \in \mathbb{R}^d ,\,
\mathbf{x} \in \mathbb{R}^{n \times d} 
$.
The constraint of the positive and negative support vectors ($x^+$ and $x^-$, respectively) satisfying
\begin{align}
	\mathbf{w} \cdot x^+ + b &= 1 \\
	\mathbf{w} \cdot x^- + b &= -1 
\end{align}
can be generalized to 
\begin{equation}
	\label{eq:svm01}
	y_i(\mathbf{w} \cdot \mathbf{x}_i + b ) \geqslant 1 %\qquad i=1,\dots,n
\end{equation}
where $y_i$ being the labels of the training data. 
The width of the margin can be inferred by projecting the vector spanning between two support vectors 
on opposite sides of the decision boundary onto the unit vector perpendicular to the hyperplane. 
If we now take a vector from a positive support vector $x^+$ to a negative support vector $x^-$ and project it on to the unit vector of $w$ (which is perpendicular to the hyperplane), 
we get the width of the margin. 
\iffalse
\begin{equation}
	\label{eq:svm02}
	\begin{aligned} 
		2\tau &= (x^+ - x^-) \cdot \frac{\mathbf{w}}{\|\mathbf{w}\|} \\
			&= (x^+ \cdot \mathbf{w} - x^-\cdot \mathbf{w} ) \cdot \frac{1}{\|\mathbf{w}\|} \\
			&= ( (1-b) - (-1-b) ) \cdot \frac{1}{\|\mathbf{w}\|} \\
			&= \frac{2}{\|w\|}
	\end{aligned}
\end{equation}
\fi
\begin{equation}
	\label{eq:svm02}
	\begin{aligned} 
		2\tau &= (x^+ - x^-) \cdot \frac{\mathbf{w}}{\|\mathbf{w}\|} 
			= (x^+ \cdot \mathbf{w} - x^-\cdot \mathbf{w} ) \cdot \frac{1}{\|\mathbf{w}\|} \\
			&= ( (1-b) - (-1-b) ) \cdot \frac{1}{\|\mathbf{w}\|} 
			= \frac{2}{\|\mathbf{w}\|}
	\end{aligned}
\end{equation}
%\td{draw sketch and make \ref{eq:svm02} on two lines}
Thus maximizing the margin is equivalent with minimizing $\|\mathbf{w}\|$ and minimizing $\frac{1}{2}\|\mathbf{w}\|^2$.
By incorporation the contraint (eq. \ref{eq:svm01}) via the Lagrangian multiplier method we arrive at the loss function: 
\begin{equation}
	\label{eq:svm03}
	\mathcal{L} = \frac{1}{2} \|\mathbf{w}\| - \sum_i^n \alpha_i [ y_i ( w_i \cdot x_i + b) -1 ] 
\end{equation}
which can be rewritten by some mathematical wizardry (setting the partial derivatives of the Lagrangian function
$\frac{\partial \mathcal{L}}{\partial \mathbf{w}}$ and $\frac{\partial \mathcal{L}}{\partial b}$ 
to zero and inserting into eq. (\ref{eq:svm03}))\cite{winston1992artificial,cherkassky1998learning}: 
\begin{equation}
	\mathcal{L} = \sum_i^n \alpha_i - \frac{1}{2} \sum_i^n \sum_j^n \alpha_i \alpha_j y_i y_j \mathbf{x}_i \cdot \mathbf{x}_j 
\end{equation}
%
\Gls{svm}s can handle non-perfect data by introducing a penalization for wrongly categorized samples. 
\begin{equation}
    L_H = sum_i^n max(0, 1- y_i ( \mathbf{w} \mathbf{x}_i + b ) 
\end{equation}
if the prediction and the true category don't agree, they have opposite signs and 
their product will be a negative number.
This is called a soft-margin \gls{svm} in contrast to the previously hard-margin \gls{svm}.

A big advantage of \gls{svm} is that through the kernel trick non-linearly separable data 
can be divided. 
\td{write about kernel trick and non strick}
